\chapter{Vertical relation: Upstream (M) and Dowstream (R)}
\section{Introduction}
\paragraph{Vertical relation}
Relationship between upstream and downstream parties: part supplier and manufacturer, manufacturer and retailer, franchiser and franchisee \emph{etc.} Whether a party is upstream or downstream depends on whether it is \emph{closer} to the consumer. For example: 
\begin{itemize}
    \item Brick-and-mortar store: supplier are upstream
    \item Online platforms: supplier are downstream
\end{itemize}
\paragraph{Vertical restraint}
Agreements between upstream and downstream parties: supply contracts, distribution contracts, licensing. Agreement/contract can serve different purposes. For example
\begin{itemize}
    \item Revenue sharing/pricing: linear pricing, \emph{non-linear} tariffs (\emph{e.g.} two part tariff). They can have a variety of names.
    \begin{itemize}
        \item Franchise fees
        \item Quantity rebates
        \item Fidelity rebates
        \item Slotting fees: manufacturer pays the retailer first.
        \item Menus of pricing schemes
    \end{itemize}
    \item Obligations/behavior restriction
    \begin{itemize}
        \item Service: M requires R provide x services to consumers.
        \item Quantity: quotas
        \item Price: Resale Price Maintenance (RPM), price ceiling/floor
        \item Bundling
        \item \textbf{Exclusivity}: exclusive territory, exclusive dealing
    \end{itemize}
    \item Integration/Merge and Acquisition: \emph{vertical} merge rather than horizontal.
\end{itemize}
\subsection{Antitrust Law}
\paragraph{Sherman Act}
\begin{enumerate}
    \item Section 1: prohibits contracts (or conspiracies) that restrain trade or commerce
    \item Section 2: prohibits monopolization or attempts to monopolize
\end{enumerate}
\paragraph{Treaty of Rome}
\begin{enumerate}
    \item Article 101: restricting competition is forbidden
    \item Article 102: prohibits abuses of a dominant position
    \begin{itemize}
        \item exploitative abuses: unfair terms (excessive prices, discrim.)
        \item exclusionary abuses: distorting competition (vert. foreclosure)
    \end{itemize}
\end{enumerate}

\section{Vertical coordination}
In this section, we will look at two coordination problems between M and R.
\subsection{Double marginalization}
\paragraph{Modeling choice}



\paragraph{Solution concept and algorithm} (Subgame perfect) Nash equilibrium using backward induction.

\paragraph{Result} R sets retail price p higher than the joint profit maximizing (\textit{jointly optimal}) retail price $p^M$.

\paragraph{Policy} There are several solutions to the problem. Each has different instrument, for example, margin, price, quantity.
\begin{enumerate}
    \item Eliminate one \textbf{margin}: \emph{e.g.} introduce more R (remove downstream margin), M will set the jointly optimal price.)
    \item Forcibly set $p^*$ as the \textbf{retail price} ceiling \footnote{The issue is that it requires M to have sufficient knowledge of the retail market.}
    \item Contracting on \textbf{retail quantity}: impose a minimum sales target on R.
    \item Align \textbf{incentive}: implement two-part tariff $\pa{w^*=c, F^*}$ \footnote{The fixed fee is used to share the profits.}
\end{enumerate}

\paragraph{Implication} The industry has incentive to remove double marginalization so as to maximize joint profit. The outcome is also good for the consumer. In this case, authority should \textit{laissez-faire}.

\subsection{Retail effort issue}
\paragraph{Modeling choice}
\paragraph{Result} Not only does R sets retail price $p>p^M$, but also it provides effort $e<e^M$.
\paragraph{Policy} Notice that on the demand side, \textbf{quantity} depends on \textbf{price} and \textbf{effort}. 
\begin{enumerate}
    \item Contract on two out of the \textbf{three dimensions}.
    \item Align \textbf{incentive} retailer becomes the residual (\textit{total profit - fixed fee}) claimant.\footnote{Increase the number of targets doesn't necessarily require additional instruments.}
\end{enumerate}
\paragraph{Question} Can we also introduce intraband competition as in double marginalization? Answer: it doesn't generate the jointly optimal $p^M$ and $e^M$. The effect can be analysed on two extreme cases, depending on whether other Rs can freeride one retailer's effort.
\begin{itemize}
    \item Free riding: effort is lower than $e^M$
    \item No free riding: effort can be either too low or too high. 
\end{itemize}
\subsection{Discussion}


%%%
\section{Horizontal coordination}

\subsection{Upstream coordination}
\paragraph{Modeling}
There are $M_A$, $M_B$ selling differentiated product to retailer $R_1$.
The first best result is such that as if $M_A$ and $M_B$ are competing directly on the market. \footnote{For example, the simplest Bertrand competition that we learn from undergrad econ.} Here, we analyse two cases. The case where seller $M$ has the bargaining power/makes the first move is more interesting. 
\paragraph{Result} Similar as before, double marginalization.
\paragraph{Solution} Following from the discussion in previous sections, in order to achieve industry-wide optimal outcome, we would need two-part tariff to align the incentive. 
\paragraph{Discussion} \label{upstr_coor:discussion} With incentive aligned, there are three potential outcomes.
\begin{itemize}
    \item $\pa{M_A, R}$ exclusive dealing:
    \item $\pa{M_B, R}$ exclusive dealing:
    \item $\pa{M_A, M_A, R}$ common agency: 
\end{itemize}
\begin{remark}
    In the absence of agency problems, there always exists a (more efficient) equilibrium with common agency.
\end{remark}
\begin{remark}
    There's always ``\textit{agency rent}'' left to the retailer. Manufacturers can not appropriate all industry-wide profit.
\end{remark}
\begin{remark}
    Public or secrete contracting doesn't matter.
\end{remark}

\subsection{Downstream coordination}
There are one upstream $M$ and two downstream firms $R_1,R_2$. Let us suppose that $M$ has the bargaining power. 
\begin{remark}
    There exhibits differences between public contracting and secrete contracting.
\end{remark}
\subsection*{Public contract}
\paragraph{Modelling: M power}
$M$ wants to set $\pa{w_1,w_2}$ such that the market outcome is the monopoly differentiated/identical one.
\paragraph{Result} If $M$ charge at costs as solving the double marginalization problem, then due to downstream competition, the price will be lower than monopoly price.
\paragraph{Policy} Let $M$ charge \textbf{above costs}, less or equal to \textbf{below monopoly price}.

\paragraph{Modelling: S power}
Retailers offer tariff to $M$. 
\paragraph{Policy} We introduce \textit{three-part tariff}--slotting fee. Each $R_i$ charges $M$ a certain amount of slotting fee $S_i$ and set $w_i=c_i$. If $R_i$ is the only retailer in the market, the industry $\pa{M,R_i}$ total profit is $\pi^M$. So it promised $M$ that $F_i=\pi^M$ will be given back as long as it is unique. If the $M$ agrees, then in the end $R_i$ gets $S_i$ and $Manufacturer$ gets $\pi^M-S_i$.
\begin{remark}
    This is \emph{de facto} exclusive dealing.
\end{remark}
Since the two retailer produces differentiated product, they can have different monopoly profit $\pi_i^M$. The more efficient retailer with higher $\pi_i^M$ wins the exclusive deal. 
\paragraph{Discussion}
Unlike the equilibrium discussed in \ref{upstr_coor:discussion}, with the slotting fee introduced above, only exclusive dealing equilibrium exists. Because both $M$ and $R$ could benefit from exclusive dealing.

\subsection*{Secrete contract}
\paragraph{Modeling: M power} 
There is \textbf{upstream opportunism} when $R_i$, $R_j$ do not observe the contract received by the other. When $M$ is contracting with one retailer, it can free-rides on other retailer’s sales.
\paragraph{Result} We discuss the outcome under Cournot and Bertrand competition.
\begin{remark}
    Depending on the beliefs of the retailer (\textit{``How to interpret deviant offers from $M$''}), there exists different equilibrium.
\end{remark}
\begin{itemize}
    \item Cournot competition
    \begin{itemize}
        \item Passive belief: perfect competition outcome
        \item Symmetric belief: monopoly outcome
    \end{itemize}
    \item Bertrand competition
    \begin{itemize}
        \item Passive belief: Direct Bertrand competition
    \end{itemize}
\end{itemize}
\paragraph{Policy}
\begin{itemize}
    \item Reputation
    \item Exclusive dealing
    \item RPM: 
        \begin{quote}
            industry-wide price floor\\
            bilateral price ceilings to squeeze downstream margin
        \end{quote}
\end{itemize}

\paragraph{Implication}
There are conflicting interests between firms’ profits and consumer surplus or total welfare. There's ground for intervention.
%%% 
\section{Competing vertical structures}
\subsection{Strategic delegation/Competition among vertical structures}
\paragraph{Modeling}
insert a picture.
There are two $M_A, M_B$. $M_A$ deals with $R_1^A, R_2^A$. Retailer only resells, therefore each retailer sells exactly the same product. Due perfect competition, retailers earn zero margin. For M, setting whole sale $w_i^j$ amounts to setting retail price. It is as if brand A and B are directly competing with each other.
We regard this scenario as the \textbf{benchmark}.
\paragraph{Policy} Now $M_A$ deals exclusively with $R_1^A$.
\paragraph{Implication} The implementation of exclusive territory dampens competition (strategic delegation). It reduces intraband competition. Retail prices respond to rival’s wholesale prices, whih gives rise to higher wholesale (and retail) prices.
\begin{remark}
    clear conflict between private/social interests
\end{remark}
\paragraph{Discussion}
Issue of observability and credibility. 
Credibility differs in types of competition.
\begin{quote}
    Bertrand competition: strategic complementarity\\
    Cournot competition: strategic substitute.
\end{quote}
    
\subsection{Facilitating practice}
\paragraph{Background} There exists different types of coordination/collusion. 
\begin{itemize}
    \item Tacit coordination
    \item Explicit coordination: communication, meetings, down raids, leniency programs
    \item Tacit coordination with explicit ``\textbf{facilitating practices}'': information sharing, trade associations, marketing practices (``price insurance''), multi-market contact, \textbf{vertical restraints} \emph{etc.} 
\end{itemize}
\paragraph{Modelling: Dynamic setting}
There are two M $\pa{M_A,M_B}$, each dealing with a single retailer $\pa{R^A,R^B}$. There's no intraband competition. In a static setting with deterministic demand, each adopting two part tariff will achieve Bertrand competition outcome. However, now 1) Manufacturers are long lived with some discount factor $\delta$. 2)There's demand shock to both products.
\paragraph{Policy} RPM as vertical restraint.
\begin{remark}
    Due to shock, price flexibility is desirable for the industry. There's a tradeoff between RPM (as collusion facilitating device) and price flexibility.
\end{remark}
\paragraph{Discussion}
\paragraph{Implication}
We know that RPM \emph{can} help firms collude and generate higher industry profit. We now analyze the impact of RPM on consumers. \\
When shocks are on demand: consumers like price rigidity. Whether RPM is beneficial to customers depends on the tradeoff between price rigidity and (high) price level.\\
When shock on cost: consumers prefer flexible prices. RPM reduces consumer welfare surely. 

\section{Multilateral relations}
\section{Exclusionary practices}


